---
title: "RADAr Summary"
output: html_document
date: "2023-07-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
```
## RADAr NED - Dataset
The dataset nedall is retrieved from a csv containing all RADAr NED data (currently to Oct 2023), and transformed so that the responses are numeric or NA. In this data set, each row is an item and each column is a test taker. The column and row IDs contain relevant information about each item or test taker. 
The code below retrieves the file and puts it into a data frame, preserving the row and col names.
```{r retrieve data from csv}
nedall <- read.csv("RADArNEDall.csv", row.names=1,header = TRUE)
```
What are the different values in this set? I can see from below that the values range from 0 to 8. I know that most items are coded 0 or 1 (dichotomous) and the hotspot items are coded 0-8. No values are unexpected. 
```{r what are the unique values in nedall - check}
# Apply unique() function to every column in the data frame
unique_values <- unique(unlist(lapply(nedall, unique)))
# Print the unique values as a vector
print(unique_values)
```
The set should now be purely numeric. Let's test that. The table function shows only one value - TRUE - meaning that all data is now numeric
```{r is nedall numeric?}
is_numeric <- sapply(nedall, is.numeric)
table(is_numeric)
```
Let's have a look at the data, starting with the distribution of scores. Firstly, I'd like to see the distribution of the total number of items answered. As can be seen, the total number varies from 113 (perhaps an error) to 115, depending on test form. The vast majority are 115.

``` {r how many items do students answer? max_score_distribution}
max_score_distribution <- lapply(nedall, function(x) sum(!is.na(x)))
max_score_distribution2 <- unlist(max_score_distribution)
table(max_score_distribution2)

```
It is quite likely that all LETNED19201e test takers have lower test items because the hotspot was not used this administration. 

```{r which admins had only 114 items?}
odd_one_out_2 <- names(max_score_distribution2)[max_score_distribution2 %in% c(114)]
odd_one_out_3 <- unique(substr(odd_one_out_2, 1, 12))
odd_one_out_3
length(odd_one_out_2)
odd_one_out_2
                                                
```
Further inspection of the csv files shows that for the NWINED20221e administration, one of the dictee items was eliminated from the test results, resulting in a count of 114 items (this was possibly due to an unfair element detected after test administration). As for the LETNED19201e, this test did not have a hotspot item, which accounts for the 114 count. 
```{r create admins_nedall, vector of all admin names}
nedall_col_names <- colnames(nedall)
admins_nedall <- c(unique(substr(nedall_col_names, 1, 12)))
admins_nedall
```
Now, let's go back to `r max_score_distribution`. This data set contains the max scores for each col in `r nedall`. How can we find out what the max_score_distrbution is for each value in vector admins_nedall, which contains different strings corresponding to the first 12 characters of the col names in nedall.
```{r mean_max_table - how many items @ each admin}
mean_max <- c()
for (admin in admins_nedall) {
  admin_columns <- nedall_col_names[grepl(paste0("^", admin), nedall_col_names)]
  admin_counts <- sum(!is.na(nedall[, admin_columns]))
  admin_mean <- admin_counts / length(admin_columns)
  mean_max <- c(mean_max, admin_mean)
}
mean_max_table <- data.frame(Admin = admins_nedall, Mean_Max = mean_max)
mean_max_table

``` 
The results printed show that the suspected LETNED19201e indeed has 114 (plus one 113) count. The other admins all have whole numbers, indicating that all test takers had that count. 
In the following count, we separate the individual test scores into batches x_scores by admin

```{r let's group indiv scores by admin}
nedall_col_sums <- colSums(nedall, na.rm=TRUE)
for (admin in admins_nedall) {
  x <- colnames(nedall)[grep(paste0("^", admin), colnames(nedall))]
  y <- nedall_col_sums[names(nedall_col_sums) %in% x]
  assign(paste0(admin, "_scores"), y)
}
scores_admins <- c()
for (admin in admins_nedall) {
  x <- paste0(admin, "_scores")
  scores_admins <- c(scores_admins, x)
}
``` 
Now, let's look at the histograms of all administrations. 
```{r histograms for scores x admins}
for (name in scores_admins) {
  x <- get(name)
  y <- length(x)
  hist(x, main = "" )
  title(paste0(name, "N=", y))
}
``` 
And the boxplots. 
```{r boxplots for scores x admins}
for (name in scores_admins) {
  x <- get(name)
  y <- length(x)
  boxplot(x, main = "")
  title(paste0(name, " N =", y))
}
``` 
Inspection of these reveal some outliers in some of the sets we can look at. 
```{r outliers?}
outliers_all <- c() 
for (name in scores_admins) {
  x <-get(name)
  # Calculate the interquartile range (IQR)
  q1 <- quantile(x, 0.25)
  q3 <- quantile(x, 0.75)
  iqr <- q3 - q1

# Define the lower and upper bounds for outliers
  lower_bound <- q1 - 1.5 * iqr
  upper_bound <- q3 + 1.5 * iqr

# Identify outliers
  outliers <- x[x < lower_bound | x > upper_bound]
  outliers_all<- c(outliers_all, outliers)
# Print the outliers

}
sorted_outliers <- sort (outliers_all)
print(sorted_outliers)

```
Some are quite low - e.g. 7 - this is someone who appears to have abandoned the test after the first section.
For now, I'll leave the outliers in, but this section can be returned to later on when running analyses. 
## Functions to navigate nedall with
### checkcheck(df) 
Pulls up a (temporary) excel file so you can inspect the data in data frame df.
```{r checkcheck function}
checkcheck <- function(df) {
  # Write the data frame to a CSV file
  write.csv(df, file = "checkcheck.csv")
  shell.exec("checkcheck.csv")
}

``` 
### grabitems(df,search_terms) 
Selects only those rows (item IDs) with the search terms in them. Multiple search terms: use c("x", "y") format in the search_terms field.
```{r grab items function}
grabitems <- function(df, search_terms) {
  ## this function creates a score csv by grabbing all items with search terms in name
  ## the search terms should be in "" and c("", "" ,"" ) for more than one
  ## Find rows containing the search terms
  search_items <- unique(grep(paste(search_terms, collapse = "|"), 
                              rownames(df), ignore.case = TRUE))
  # Subset the data frame by the search items
  search_df <- df[search_items, ]
  # return data frame
  return(search_df)
}  
``` 
### grabtotalscores(df, search_terms)
```{r grab total scores function}
grabtotalscores <- function(df, search_terms) {
  
  # create an empty data frame to store the results
  sumsdf <-data.frame()
  
  # loop through the search terms
  for (term in search_terms) {
    # grab the TOTAL row for the current search term
    print(term)
    grabdf <- grabitemsplus(df, term)
    grabrow <- grabdf["TOTAL",]
    sumsdf <- rbind(sumsdf, grabrow)
    
  }
  grabdf <- grabitemsplus(df, search_terms[1])
  form <- c(grabdf["FORM", ])
  admin <- c(grabdf["ADMIN", ])
  sumsdf <- rbind(sumsdf, form, admin)
  
  
  rownames(sumsdf) <- c(search_terms, "FORM", "ADMIN")
  sumsdf <- sumsdf[,1:(ncol(sumsdf)-3)]
  
  
  return(sumsdf)
}

``` 
## Question: Whole Test Analysis: What percentage of students 'pass' (70%) each form?
The pass rate for NED is 70%. What does this mean for different administrations.
Using the custom function grabitemsplus, additional information is added to nedall as nedallplus
First, we will create nedallplus, a data frame which has additional information. First we will add ADMIN as the first row
```{r  nedallplus - add row ADMIN}

# Create a new row called "ADMIN" with extracted characters from column headings
# Get the column names of the dataframe nedall

# Get the column names of the dataframe nedall
column_names <- colnames(nedall)

# Create a new row with extracted characters and name it "ADMIN"
ADMIN <- substr(column_names, 1, 12)

# Bind the new row to the top of the dataframe
nedallplus1 <- rbind(ADMIN, nedall)
rownames(nedallplus1)[1] <- "ADMIN"
```
Now I want to add another row with item numbers, derrived above with variable `r `mean_max`
```{r add row: Number Items Answered <- nedallplus2}
# Look up values and add to ITEMS vector
ITEMS <- c()
admin_vector <- as.vector(nedallplus1["ADMIN", ])

for (i in 1:ncol(nedallplus1)) {
  admin_value <- admin_vector[i]
  match_row <- match(admin_value, mean_max_table$Admin)
  mean_max_value <- mean_max_table$Mean_Max[match_row]
  ITEMS <- c(ITEMS, mean_max_value)
}
nedallplus2 <- rbind(ITEMS,nedallplus1)
rownames(nedallplus2)[1] <- "NUMBER_ITEMS"
```
Now, some of the administrations use exactly the same form, so the following chunk creates a matrix with admin in col 1 and form in col 2.
```{r create admin to form matrix for lookup purposes (forms_to_admins_vector)}
forms_to_admins_vector <- c("FTRNED21221e","FTRNED(A)",
                            "FTRNED22231e","FTRNED(A)",
                            "LETNED19201e","LETNED(A)",
                            "LETNED20211e","LETNED(B)",
                            "LETNED21221e", "LETNED(B)",
                            "LETNED22231e", "LETNED(B)",
                            "LETNED20212e", "LETNED(C)", 
                            "LETNED21222e", "LETNED(C)",
                            "MANNED20211e", "MANNED(A)",
                            "MANNED21221e", "MANNED(A)",
                            "MANNED21222e", "MANNED(A)",
                            "MANNED22231e", "MANNED(B)",
                            "MEDNED21221e", "MEDNED(A)",
                            "MEDNED22231e", "MEDNED(A)",
                            "NWINED21221e", "NWINED(A)", 
                            "NWINED22231e", "NWINED(A)",
                            "SOWNED22231e", "SOWNED(A)")
forms_admins_matrix <- matrix(forms_to_admins_vector,ncol=2,byrow=TRUE)
print(forms_admins_matrix)
```
Now I want to add another row to nedallplus2 called FORM. 
```{r add row FORM to nedallplus2}
FORM <- c()
admin_vector <- as.vector(nedallplus2["ADMIN", ])
for (i in 1:ncol(nedallplus2)) {
  admin_value <- admin_vector[i]
  match_row <- match(admin_value, forms_admins_matrix[,1])
  form_value <- forms_admins_matrix[,2][match_row]
  FORM <- c(FORM, form_value)
}
nedallplus3 <- rbind(FORM, nedallplus2)
rownames(nedallplus3)[1]<- "FORM"
``` 
The hotspot items are the only non-dichotomous items. Therefore the NUMBER_ITEMS row does not actually give us the maximum score for each form/admin. Cross-referencing all NED hotspot items in use in Cirrus confirms that all have a max of 7. in'to Languages has a way to adjust this to a max of 8, so scores in the data have a max of 8. However, one form - LETNED(A) - does not have a hotspot at all, so the NUMBER_ITEMS is acutally the max. 

In any case, to make this easier, I'll first add the hotspot max score 8 or 0 to the matrix we used earlier for forms, forms_to_admins_vector. 
```{r forms_admins_hotspots_matrix - hotspot points max by admin and form}
hotspot_max <- c(8,8,0,8,8,8,8,8,8,8,8,8,8,8,8,8,8)
forms_admins_hotspots_matrix <- cbind(forms_admins_matrix, hotspot_max)
```
Now I'll use this once again to look up the value and add it to NUMBER_ITEMS in nedallplus3 to create nedallplus4 with row MAX_SCORE
```{r nedallplus4 - new row MAX_SCORE} 
hotspot_max <- c()
admin_vector <- as.vector(nedallplus2["ADMIN", ])
for (i in 1:ncol(nedallplus3)) {
  admin_value <- admin_vector[i]
  match_row <- match(admin_value, forms_admins_hotspots_matrix[,1])
  form_value <- forms_admins_hotspots_matrix[,3][match_row]
  hotspot_max <- c(hotspot_max, form_value)
}

nedallplus4 <- rbind(hotspot_max, nedallplus3)
rownames(nedallplus4)[1]<- "HOTSPOT_MAX"
# calculate row MAX_SCORE
x <- as.vector(as.numeric(nedallplus4["HOTSPOT_MAX",]))
y <- as.vector(as.numeric(nedallplus4["NUMBER_ITEMS",]))
z <- c(x+y)
nedallplus5 <- rbind(z, nedallplus4)
rownames(nedallplus5)[1]<- "MAX_SCORE"
```
Now lets add another row for raw score
```{r RAW_SCORE added}
col_sum <- c()



for (i in 1:ncol(nedallplus5)) {
  x <- nedallplus5[6:nrow(nedallplus5),i]
  y <- as.numeric(x)
  z <- sum(y, na.rm=TRUE)
  col_sum <- c(col_sum, z)
}
nedallplus6 <- rbind(col_sum, nedallplus5)
rownames(nedallplus6)[1]<- "RAW_SCORE"
```
Now that we have the RAW_SCORE, we can calcluate the percentage score
```{r PERCENT_SCORE added}
x <- as.vector(as.numeric(nedallplus6["MAX_SCORE",])) 
y <- as.vector(as.numeric(nedallplus6["RAW_SCORE",]))
z <- c(y/x)
nedallplus7 <- rbind(z, nedallplus6)
rownames(nedallplus7)[1]<- "PERCENT_SCORE"
```
Finally, let's transform the last interation of nedallplus (currently nedallplus7) to a generic name, nedallplus
```{r - TRANSFORM to nedallplus}
nedallplus <- nedallplus7
```
`r nedallplus` contains information that will now allow us to inspect different forms for differences. We choose to use forms instead of admins because many admins have only a few tts. 
To do analyses, we need to transform the data so that the col names are the summary data we've added along with item names. 
Let's start by summarizing the Ns for each form using variable $FORM to group a summary:
```{r forms_scores: N, Mean, SD, Min, Max}
forms <- unlist(as.vector(nedallplus["FORM", ]))
scores <- as.numeric(unlist(as.vector(nedallplus["PERCENT_SCORE",])))

summary <- forms_scores %>%
  group_by(FORM) %>%
  summarise(
    count = n(),
    min_score = min(SCORES),
    max_score = max(SCORES),
    mean_score = mean(SCORES),
    sd_score = sd(SCORES)
  )

# Print the result
print(summary)
```

let's look at these in a violin histogram. The red line indicates the average for the overall test, which is stored in variable `r overall_mean`. The chart shows that the individual means for the forms are near but not on the line, indicating that some groups score better on their test form than other groups. Importantly, the test forms are different and the students are siloed, so we cannot yet say that there is an issue of unfairness. Additionally, we do not yet know if these differences are statistically significant. That is the next step. 
```{r For FORMS - Violin Distribution Chart}

# Calculate means by form
overall_mean<- forms_scores %>%
  summarize(mean_score = mean(SCORES))

overall_mean <-unlist(overall_mean)
cat("overall average:", overall_mean)


mean_scores <- forms_scores %>%
  group_by(FORM) %>%
  summarise(mean_score = mean(SCORES))

# Create the violin plot
ggplot(forms_scores, aes(x = FORM, y = SCORES)) +
  geom_violin(fill = "gray", color = "gray") +
  geom_hline(aes(yintercept = mean(SCORES)), linetype = "dashed", color = "red") +
  geom_point(data = mean_scores, aes(x = FORM, y = mean_score), color = "black", size = 3, shape = 19) +
  labs(title = "Distribution of scores in % by form") +
  xlab(NULL) +
  ylab("Scores") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```
The appearence of outliers on the lower end is quite apparent here. It is a good idea now to look back at our outlier list and determine a method by which they can be evaluated for exclusion. The variable `r outliers_all`, calculated earlier, contains the outliers: testtaker and score. Below, a new data frame, `r outliers_df` contains these two rows. But first, let's inspect the outliers raw scores. 
```{r distribution of outliers}
hist(outliers_all)
```
Which should we exclude? Firstly, those who abandon the test. To see this, we need to look at the individual responses to items to see the behaviour. Below, `r outliers_details_collapsed` provides this information. 

```{r exclude outliers? outliers_all}

names_outliers_all <- names(outliers_all)
outliers_df <- data.frame(testtaker=names_outliers_all, score=outliers_all)
outliers_details <- nedallplus_t[outliers_df$testtaker,]
# Collapse the data frame
outliers_details_collapsed <- as.data.frame(t(apply(outliers_details, 1, function(x) {
  x <- na.omit(x)
  c(x, rep(NA, ncol(outliers_details) - length(x)))
})))

# Print the collapsed data frame
print(outliers_details_collapsed)
checkcheck(outliers_details_collapsed)
```
We can use this new data frame `r outliers_df`, to capture the individual responses for items for each testtaker and use that to extend the outliers_df. 
```r{exploring outliers_df: collecting item responses}




# Create a line plot
plot(scores, type = "l", xlab = "Index", ylab = "Score", main = "Sequence of Scores")


```
Are any of these differences significant? An ANOVA will help us answer this. There are assumptions about the data that must be checked before the ANOVA is performed. Independence: The data is independent. Normality: There is some question about the normality of the distributions of scores as the tend to be left skewed due to a ceiling effect. Homogeneity of Variance: This will need to be checked as well. Random Sampling: In this case, the sample is the population, so this assumption is met. Let's check now for a normal distribution. 
```{r normal distribution of scores in forms_scores?}
results <- split(forms_scores$SCORES, forms_scores$FORM) %>%
  lapply(shapiro.test)

# Print the test results
for (group in unique(forms_scores$FORM)) {
  print(paste("FORM", group))
  print(results[[group]])
}
```

```{r}
#anova
```
Which ones? (post-hoc)
```{r}
#post-hoc procedures
```
What are the pass/fail rates for each form (70% for NED)?
```{r}
#create empty vector passfail()
#using nedallpluslong
#$FORM and $PERCENT
# for loop for form in allforms
# gather all $PERCENT for $FORM
# count number >= 70, <- pass
# total - pass = <- fail 
# pass/total <- perc_pass
# pass FORM, total, pass, fail, perc_pass to vector passfail
# create matrix passfail_matrix from vector passfail using form as row names and designate col names manually
# print(passfail_matrix)
```
Are there significant differences in the expected vs observed pass/fall numbers
```{r}
# use passfail_matrix for analysis
# use chisquare to determine significant differences
# post hoc to identify where
```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```













